---
title: "TreeLife demo"
author: "Jaideep Joshi"
date: "2023-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r}
library(PlantFATE)
library(tidyverse)

print(getwd())
```

## Calculate fitness of specified traits

To calculate the fitness of a tree with given traits, follow these steps:

1. Create an optimizer
2. set the parameters file
3. Initialize the optimizer - this will read the params file and initialize everything from it
4. set traits  - this will overwrite the specified traits and redo any trait-coordination calculations
5. Calculate fitness

```{r}
lho = new(LifeHistoryOptimizer)
lho$params_file = "tests/params/p.ini"
lho$init()
lho$set_traits(c(0.08, 750))
lho$printPlant()
lho$calcFitness()
```

## Calculate fitness as a function of LMA

```{r}
fitness_lma = function(lma){
  lho = new(LifeHistoryOptimizer)
  lho$params_file = "tests/params/p.ini"
  lho$init()
  lho$set_traits(c(lma, 750))
  # lho$printPlant()
  lho$calcFitness()
}

lma = seq(0.05, 0.3, 0.01)
dat = lma %>% purrr::map_dbl(fitness_lma)

plot(dat~lma)
```

## Calculate fitness as a function of wood density

```{r}
fitness_wd = function(wd){
  lho = new(LifeHistoryOptimizer)
  lho$params_file = "tests/params/p.ini"
  lho$init()
  lho$set_traits(c(0.12, wd))
  # lho$printPlant()
  lho$calcFitness()
}

wd = seq(350, 900, length.out=20)
dat = wd %>% purrr::map_dbl(fitness_wd)

plot(dat~wd)
```
